---
title: "BBS Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r, echo=FALSE}

library(tidyverse)
library(dplyr)
library(FD)
library(spData)
library(sf)
library(tmap)
library(furrr)


#piggyback::pb_download(repo = "karinorman/functional_diversity")

load("~/Documents/Projects/functional_diversity/data/trait.rda")
load("~/Documents/Projects/functional_diversity/data/bbs.rda")


min_year = 2006 #define the minimum year of sampling to include

p <- st_crs(wkt='PROJCS["USA_Contiguous_Albers_Equal_Area_Conic",
    GEOGCS["GCS_North_American_1983",
            DATUM["North_American_Datum_1983",
            SPHEROID["GRS_1980",6378137,298.257222101]],
            PRIMEM["Greenwich",0],
            UNIT["Degree",0.017453292519943295]],
            PROJECTION["Albers_Conic_Equal_Area"],
            PARAMETER["False_Easting",0],
            PARAMETER["False_Northing",0],
            PARAMETER["longitude_of_center",-96],
            PARAMETER["Standard_Parallel_1",29.5],
            PARAMETER["Standard_Parallel_2",45.5],
            PARAMETER["latitude_of_center",37.5],
            UNIT["Meter",1],
            AUTHORITY["EPSG","102003"]]')

```
We want to be able to figure out if a site's Functional Diversity (FD) is significantly different than what we would expect for a site with that richness level in that region. First we must simulate a null model of the relationship between FD and species diversity for a given region.

Code below attempts to create a null model curve (n = 1) for the Northern Rockies.
```{r}
#site level FD with region assignments
bbs_site_FD <- get_complete_site_data()

#function to calculate FD from a list of species
get_sample_fd <- function(x, richness, species_pool){
  samp_trait_mat <- get_trait_matrix(sample(species_pool$scientific, richness)) #sample needs list, so specify the column name to change type 
  sample_FD <- dbFD(x = samp_trait_mat, w.abun = FALSE)
  return(head(sample_FD, -1)) #remove last element, which is the CWM for each trait
}

#get sample summary statistics for each richness level observed in a region 
simulate_null <- function(region_data, n = 1000){
  
  species_pool <- bbs %>% 
    filter(year > min_year & site_id %in% unique(region_data$site_id)) %>%
    select(scientific) %>%
    unique() #get column of unique scientific names

  richness_levels <- unique(region_data$nbsp)
  
  #Get n FD samples for a given richness value
  null_samples <- function(richness){
    fds <- map_dfr(1:n, possibly(get_sample_fd, data.frame()), richness = richness, species_pool = species_pool)
    
    if(dim(fds)[1] > 0){
      print("yes")
      fds <-  fds %>%
        drop_na() %>% #after parallelizing, ocassionally one sample will result in NA for FDiv, still trying to figure out why
        select(-nbsp, -sing.sp) %>%
        gather(factor_key = TRUE) %>%
        group_by(key) %>%
        summarise(mean = mean(value), sd = sd(value)) %>%
        mutate(se = sd/sqrt(n),
               lowerCI = mean - qt(0.975, n - 1) * se, 
               upperCI = mean + qt(0.975, n - 1) * se) %>%
        # gather(variable, value, -key) %>%
        # unite(temp, key, variable) %>%
        # spread(temp, value) %>%
        mutate(richness = richness) %>%
        rename(metric = key) %>%
        select(richness, everything())
    }
    return(fds)
  }
  
  test_sim <- future_map_dfr(richness_levels, null_samples) #should be future_map
}

###Simulate null model for one region###
get_regional_sig_test <- function(region_name, ...){
  
  #get sites in the Northern rockies region and get species pool for that region
  region_FD <- dplyr::filter(bbs_site_FD, region == region_name)

  plan(multicore)
  null_stats <- simulate_null(region_FD, ...)

  write_tsv(null_stats, paste0("~/functional_diversity/data/stat/", region_name, ".tsv.bz2"))
  
  metric_names <- c("FRic", "qual.FRic", "FEve", "FDiv", "FDis", "RaoQ")
  
  FD_sig <- region_FD %>%
    select(-sing.sp, -starts_with("CWM")) %>%
    gather(metric_names, key = "metric", value = "value") %>%
    left_join(null_stats, by = c("nbsp" = "richness", "metric")) %>%
    mutate(significant = ifelse(value > lowerCI & value < upperCI, FALSE, TRUE)) %>%
    select(-c("value","region", "mean", "sd", "se", "lowerCI", "upperCI")) %>%
    spread(metric, significant)
  
  write_tsv(FD_sig, paste0("~/functional_diversity/data/stat_test/", region_name, ".tsv.bz2"))
  
  return(TRUE)
}

get_metric_sig <- function(){
  if(!dir.exists("~/functional_diversity/data/stat")){
    
    dir.create("~/functional_diversity/data/stat")
    dir.create("~/functional_diversity/data/stat_test")
    
    out <- map(unique(bbs_site_FD$region), get_regional_sig_test)
  }
  
  stat_test <- purrr::map_dfr(fs::dir_ls(path = "~/functional_diversity/data/stat_test/", glob="*.tsv.bz2"), readr::read_tsv, .id = "region")
}

stat_test <- get_metric_sig()

#n_rockies <- get_regional_sig_test(region_name = "CENTRAL HARDWOODS", n = 10)
stats <- purrr::map_dfr(fs::dir_ls(path = "~/functional_diversity/data/stat/", glob="*.tsv.bz2"), readr::read_tsv, .id = "region")

```

```{r}
#preliminary plot of null curve
n_rockies_null %>% 
  #filter(metric == "FRic") %>%
  ggplot(aes(x = richness, y = mean)) + geom_smooth() +
  theme_classic() +
  facet_grid(.~metric)
```

Map of regions with bbs sites
```{r}
bcr <- get_ecoreg_shp()

#dissolve region polygons to get rid of state/providence boundaries
bcr_regions <- lwgeom::st_make_valid(bcr) %>%
  #filter(COUNTRY %in% c("USA","CANADA")) %>%
  st_set_precision(-10000) %>%
  group_by(BCRNAME) %>%
  summarize()

#get basemap
data("World")
nam <- st_as_sf(World) %>% 
  filter(continent == "North America", subregion == "Northern America")
st_crs(nam) <- 54012

#map
regions_ma <- tm_shape(nam, projection = p$proj4string) +
  tm_borders() +
tm_shape(bcr_regions) +
  tm_borders("white") +
  tm_fill(col = "gold2", alpha = 0.25) +
tm_shape(st_as_sf(bbs_site_FD)) +
  tm_dots()

regions_ma
```

